---
title: "Tarea5_Cap6"
author: "Alexandra Lopez"
date: "2025-04-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
library(discrtr) # A companion package for the book Introduction to Discrete Choice Analysis with `R`
library(dplyr) # A Grammar of Data Manipulation
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(mlogit) # Multinomial Logit Models
library(readr) # Read Rectangular Text Data
library(gplots) # Various R Programming Tools for Plotting Data
library(ggmosaic) # Mosaic Plots in the 'ggplot2' Framework
library(treemapify)
library(ggridges)
library(ggalluvial)
library(evd)
library(htmlwidgets) # HTML Widgets for R
library(kableExtra) # Construct Complex Table with kable and Pipe Syntax
library(plotly) # Create interactive web graphics
library(tidyr) # Tidy messy data
#library(webshot2) # Take screenshots of web pages

library(evd)
```



*1. What is the difference between a marginal effect and an elasticity?*

Los efectos marginales y las elasticidades son formas de cuantificar el impacto de un cambio en una variable independiente sobre el resultado de un modelo, pero difieren en cómo expresan ese cambio.

El efecto marginal mide el cambio absoluto en la probabilidad (o variable de resultado) cuando una variable independiente cambia en una unidad, manteniendo todo lo demás constante.

La elasticidad mide el cambio porcentual en la variable dependiente (como la probabilidad o la utilidad) resultante de un cambio del 1% en una variable independiente.


*2. Why is it inappropriate to calculate the elasticity of a dummy variable?*

Una variable ficticia (o dummy variable) es una variable binaria que solo puede tomar dos valores (0 o 1).

La elasticidad mide: El cambio porcentual en una variable dependiente $Y$, cuando una variable independiente $X$ cambia en 1%.

Pero para una variable ficticia, esto no tiene sentido porque:

-No puedes cambiar un $0$ o un $1$ en un $1%$.

-No existe un concepto de “cambio porcentual” en algo que solo es *$sí$* o *$no$*.


*3. Use Model 3 in this chapter and calculate the marginal effects and the elasticities for operating cost at the mean of all variables.*

```{r}
data("Heating")

```


```{r}
Proportion <- Heating %>% group_by(depvar) %>% summarise(no_rows = n())

df <- data.frame(System = c("Gas Central", "Gas Room", "Electric Central", "Electric Room", "Heat Pump"),
                 Operation = c(mean(Heating$oc.gc), mean(Heating$oc.gr),
                               mean(Heating$oc.ec), mean(Heating$oc.er),
                               mean(Heating$oc.hp)),
                 Proportion = Proportion$no_rows/900)

df %>% kable() %>% kable_styling()
```


```{r}
H <- Heating %>% mlogit.data(shape = "wide", choice="depvar", varying=c(3:12))
```


```{r}

model3_Long <- mlogit(depvar~oc, H, reflevel = "ec")
summary(model3_Long)

```

```{r}

oc_mean <- Heating[1, ]
mean_cost <- Heating %>% select(starts_with("oc")) %>%
             summarise(across(.cols = everything(), mean))

oc_mean[8:12] <- mean_cost
head(oc_mean)
```

```{r}
p_model1_ic_mean <- predict(model3, newdata = mlogit.data(oc_mean, shape = "wide", choice = "depvar", varying = 8:12))
```

EFECTOS MARGINALES:

```{r}
effects(model3, covariate = "oc", type = "ra", data = mlogit.data(oc_mean, shape = "wide", choice = "depvar", varying = 8:12))
```

ELASTICIDADES:

```{r}
effects(model3, covariate = "oc", type = "rr", data = mlogit.data(oc_mean, shape = "wide", choice = "depvar", varying = 8:12))
```

*4. Use Model 3 in this chapter to calculate the rebate needed to reach a 10% penetration rate of heat pumps.*


```{r}
H_rebate <- Heating %>% mutate(oc.hp=0.90*oc.hp)
```

```{r}
data.frame(Policy = c("Do nothing", "10% rebate"),
           rbind(apply(predict(model3, newdata = mlogit.data(Heating, shape = "wide", choice = "depvar",
                       varying = c(8:12))), 2, mean),
                 apply(predict(model3, newdata = mlogit.data(H_rebate, shape = "wide", choice = "depvar",
                       varying = c(8:12))), 2, mean)))
```


*Estimate a new model that extends Model 3 by introducing the age of the household head. Use the electric room system (“er”) as the reference level.*


```{r}
model3_1 <- mlogit(depvar ~ ic + oc | agehed, Heating, shape = "wide", choice = "depvar", reflevel = "er", varying = c(3:12))

summary(model3_1)
```

*5. Use the likelihood ratio test to compare your new model to Model 3.*

```{r}
lrtest(model3, model3_1)
```


*6. Is the ratio of the coefficient of installation (or operation) cost to the coefficient of age of household head meaningful? Explain.*

Cada coeficiente debe interpretarse por separado, en relación con su propia unidad y efecto marginal sobre la probabilidad de elección.
