---
title: "TAREA_6"
author: "Alexandra Lopez"
date: "2025-05-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
library(discrtr) # A companion package for the book Introduction to Discrete Choice Analysis with `R`
library(dplyr) # A Grammar of Data Manipulation
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(mlogit) # Multinomial Logit Models
library(readr) # Read Rectangular Text Data
library(gplots) # Various R Programming Tools for Plotting Data
library(ggmosaic) # Mosaic Plots in the 'ggplot2' Framework
library(treemapify)
library(ggridges)
library(ggalluvial)
library(evd)
library(htmlwidgets) # HTML Widgets for R
library(kableExtra) # Construct Complex Table with kable and Pipe Syntax
library(plotly) # Create interactive web graphics
library(tidyr) # Tidy messy data

library(evd)
```



**2. Use model nl2 in this chapter and calculate the direct-point elasticity at the mean values of the variables, for an increase in the installation costs of Gas Central systems**

**3. Use model nl2 in this chapter and calculate the cross-point elasticity at the mean values of the variables, for a 1% increase in the operation costs of Gas Central systems.**



```{r}
data("Heating", package = "mlogit")
H <- mlogit.data(Heating, shape = "wide", choice = "depvar", varying = c(3:12))
```


```{r}
X_mean <- model.matrix(nl2)[1:5,]
alt <- index(H)$alt[1:5]
```


```{r}
nl2 <- mlogit(
  depvar ~ ic + oc, 
  data = H,
  nests = list(room = c("er", "gr"), central = c("ec", "gc", "hp")),
  un.nest.el = TRUE,
  steptol = 1e-12
)
```


```{r}
gc_data <- subset(H, alt == "gc")
gc_ic_mean <- mean(gc_data$ic, na.rm = TRUE)
```

```{r}
# Predicciones del modelo
probs <- fitted(nl2, type = "probabilities")
mean_prob_gc <- mean(probs[H$alt == "gc"], na.rm = TRUE)

# Coeficiente del costo de instalación (ic)
beta_ic <- coef(nl2)["ic"]

# Elasticidad directa para gas central
elasticity_gc_ic <- beta_ic * gc_ic_mean * (1 - mean_prob_gc)
elasticity_gc_ic

```

```{r}
gc_oc_mean <- mean(H$oc[H$alt == "gc"], na.rm = TRUE)
beta_oc <- coef(nl2)["oc"]

# Supón que queremos el efecto en 'ec' (electric central)
mean_prob_ec <- mean(probs[H$alt == "ec"], na.rm = TRUE)
elasticity_ec_gc_oc <- beta_oc * gc_oc_mean * (-mean_prob_ec)
elasticity_ec_gc_oc

```


**4. Re-estimate the nested logit model in this chapter, but change the nests to types of energy as follows:**

**• Gas: gas central, gas room.**

**• Electricity: electric central, electric room, heat pump.**

**Use a single coefficient for the inclusive variables (i.e., set un.nest.el = TRUE). Are the results reasonable? Discuss.**


```{r}
nests_energy <- list(
  gas = c("gc", "gr"),
  electric = c("ec", "er", "hp")
)

nl_energy <- mlogit(
  depvar ~ ic + oc, 
  data = H,
  nests = nests_energy,
  un.nest.el = TRUE
)

summary(nl_energy)

```



El modelo logit anidado fue ajustado con éxito y converge. 

Las alternativas tienen diferentes frecuencias de elección, siendo gc la más popular. 

Las variables ic y oc tienen un efecto negativo y estadísticamente significativo en la utilidad de las alternativas.

oc es el predictor más fuerte y significativo. 

El parámetro iv también muestra un efecto positivo marginalmente significativo. Aunque el modelo es globalmente significativo (según la prueba de razón de verosimilitudes), el bajo R-cuadrado de McFadden sugiere que hay mucha variabilidad en las elecciones que no es explicada por las variables incluidas en el modelo. Esto podría indicar que faltan otras variables importantes, que la relación es más compleja o que hay mucha variabilidad individual no observada.


